---
title: "lab8-regression"
author: "Eleonore Durand"
date: "November 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
salmon <- read_csv("salmon_fresh.csv") %>% 
  select(year, salmon_fresh)

```

Exploration of the data :

```{r}

salmon_scatter <- ggplot(salmon, aes(x=year, y=salmon_fresh)) +
  geom_point() +
  theme_classic()

salmon_scatter

#overall there does seem to be a linear relationship between the year and the amount of salmon imported (lbs), but there seems to be two clusters of data (6 points total, that dont fit the model)

```


Now we do the linear regression:

```{r}
salmon_model <- lm(salmon_fresh ~ year, data = salmon)
salmon_model


#imports (in thousands of lbs)= -14982940 +7550(year) +error
```

• What does the slope mean in the context of this model?
For every extra yr, the US imports goes up by 7550 thousands of lbs of salmon. 
• What does the y-intercept mean in the context of this model? Why is that concerning? What does this
mean about extrapolating this model for past values?
The y intercept means that for year 0, the US imported -14,982,940 thousands lbs of salmon. This is concerning because it doesn't make sense to import negative amounts of salmon in year 0. This means that you cannot use this model to extrapolate estimates before a certain year. 

Model diagnostics :

```{r}
par(mfrow = c(2,2))
plot(salmon_model)
```

Do residuals appear normally distributed? 
Any concerns about heteroscedasticity or outliers?

The residuals appear to be normally distributed according to the residuals vs fitted and the qq plots. However if there are three outlier points that are concerning/warrant a second look. 

